% Using template for a Computer Science Tripos Part II project dissertation
\documentclass[12pt,a4paper,twoside,openright]{report}
\usepackage[pdfborder={0 0 0}]{hyperref}    % turns references into hyperlinks
\usepackage[margin=25mm]{geometry}  % adjusts page layout
\usepackage{graphicx}  % allows inclusion of PDF, PNG and JPG images
\usepackage{verbatim}
\usepackage{docmute}   % only needed to allow inclusion of proposal.tex
\usepackage{natbib}

\raggedbottom                           % try to avoid widows and orphans
\sloppy
\frenchspacing
\clubpenalty1000%
\widowpenalty1000%
\parindent 0pt

\renewcommand{\baselinestretch}{1.1}    % adjust line spacing to make
                                        % more readable

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title


\pagestyle{empty}

\rightline{\LARGE \textbf{Z\'ebulon Goriely}}

\vspace*{60mm}
\begin{center}
\Huge
\textbf{Simulating Language Learning and Evolution} \\[5mm]
Computer Science Tripos -- Part II \\[5mm]
Queens' College \\[5mm]
\today  % today's date
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Proforma, table of contents and list of figures

\pagestyle{plain}

\chapter*{Proforma}

{\large
\begin{tabular}{ll}
Name:               & \bf Z\'ebulon Goriely                       \\
College:            & \bf Queens' College                     \\
Project Title:      & \bf Simulating Language Learning and Evolution \\
Examination:        & \bf Computer Science Tripos -- Part II, May 2020  \\
Word Count:         & \bf NULL\footnotemark[1]  \\
Project Originator: & Z\'ebulon Goriely                    \\
Supervisor:         & Prof. Paula Buttery and Dr. Andrew Caines                  \\ 
\end{tabular}
}
\footnotetext[1]{This word count was computed
by \texttt{detex diss.tex | tr -cd '0-9A-Za-z $\tt\backslash$n' | wc -w}
}
\stepcounter{footnote}


\section*{Original Aims of the Project}

- max 100 words

- Investigate evolutionary advantage of language

- Do this by implementing a simulation of a population of neural nets in a toy mushroom world

- Compare the fitness of three species; one without language, one with an evolved language and one with an externally imposed language 


Language has evolved and therefore probably gave an evolutionary advantage to the individuals that exhibited it. According to \citet{Cangelosi1998}, it is difficult to investigate the evolutionary origin of language and the selective pressures that may have originated language due to the limited evidence available. They propose using computer simulations of evolutionary scenarios to investigate this. In the paper referenced, they describe a simulated toy world where agents controlled by neural networks interact with an environment of mushrooms that are edible and poisonous. Exploring this simulation is the basis of my project.

\section*{Work Completed}

- max 100 words

- Implemented the simulation capable of hosting the three population types

- Genetic algorithm, feed-forward neural networks

- Graphing etc.


I implemented the mushroom-world simulation and explored three populations of feed-forward neural networks; one without language, one with an externally imposed language and one with an evolved language. To simulate evolution, I implemented a genetic algorithm to allow the fittest members of each generation to reproduce. The state of the simulation was saved at each generation in order plot the fitness of the three populations, plot the quality of the language produced and investigate behavioural tests. To conclude the project, I compared these findings with \citet{Cangelosi1998}.

\section*{Special Difficulties}

None.
 
\newpage
\section*{Declaration}

I, Z\'ebulon Goriely of Queens' College, being a candidate for Part II of the Computer
Science Tripos, hereby declare
that this dissertation and the work described in it are my own work,
unaided except as may be specified below, and that the dissertation
does not contain material that has already been used to any substantial
extent for a comparable purpose.

\bigskip
\leftline{Signed Z\'ebulon Goriely}

\medskip
\leftline{Date [date]}

\tableofcontents

\listoffigures

\newpage
\section*{Acknowledgements}

Paula et al.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Introduction

\pagestyle{headings}

\chapter{Introduction}

\emph{Introduction here, start again as if abstract does not exist. Include your contributions in bullet point format. End with paragraph describing structure of the rest of the paper}

\section{Motivations}

\emph{High level, still cite, why does this interest us from a scientific perspective. What could we learn.}

- Want to investigate evolutionary origin of language

- Difficult to investigate due to limited evidence, so computer simulations could be good

- Investigation of cultural vs sexual mechanisms

- Artificial life

- Main question - how can language evolve when it has a purely informative function and so advantageous to receiver but not to the producer?

- Reference Clark (1993) for what an efficient language is

\section{Prior Work}

\emph{State what similar work has been done before and since what i'm trying to replicate. Also a continuation of the motivation. Why does this paper in particular interest you. Focus more on their prior work. Talk about work that has continued.}

- Symbol grounding

- Language games

- Mushroom world

\section{Project Overview}

%%%%%%%%%%%%%%%%%%%%%
% Preparation

\chapter{Preparation}

\section{Mushroom World}

\subsection{Mushrooms}

- Talk about how mushrooms are represented

- Differences between edible and poisonous mushrooms

\subsection{Environment}

- Talk about size of the world, number of mushrooms in the world

- Number of epochs, cycles

\section{Entities}

- Entities represent animals that eat edible mushrooms

\subsection{Feed-Forward Neural Networks}

- Description of what a neural network is

- Description of feeding forward

- Description of structure used for these entities

\subsection{Genetic Algorithm}

- Description of algorithm

- Specific values used for this simulation

- Represents evolution, natural selection, etc.

\subsection{Population types}

- No language

- External language

- Evolved language

\section{Evaluation Metrics}

\subsection{Generational Fitness}

- 'Fitness' described as average score of all entities for each generation where score is calculated using total number of edible and poisonous mushrooms eaten by the entity

- Plot fitness across 1000 generations for each of the three population types

\subsection{Efficiency of Language}

- 'Naming experiment' to graph the language used for edible and poisonous mushrooms

- Describe how QI encapsulates the 'efficiency' of a language

- Talk about how we can investigate how language production is linked to cognitive ability to differentiate between mushrooms by looking at language produced by non-speaking populations

- Should measure correlation between QI and fitness

\section{Requirements Analysis}

My project reimplements \citet{Cangelosi1998} soooo
As such, the requirements of this project are:

- Implement simulation environment (mushrooms, cycles,)

- Implement neural-network entities with genetic algorithm, three populations

- Plot fitness over time

- Plot distributions of signals used for edible and poisonous mushrooms

- Calculate QI over time and investigate correlation with fitness 

\section{Starting Point}

- coded from scratch

- familiar with neural networks from AI course, language theory from Formal Models of Language

- familiarised self with genetic algorithm

\section{Software Engineering}

\subsection{Languages and Libraries}

python, numpy, pytest

\subsection{Project Management}

- Agile

- How I stuck to project plan

\subsection{Version Control}

- github, hardware backups

\subsection{Development tools}

travis, pylint, yapf

\subsection{Development environment}

VSCode, Slurm w/ HPC 

\subsection{Code License}

MIT?

\section{Summary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Implementation

\chapter{Implementation}

\emph{Start with UML, high level overview}

\emph{Talk about problems with trying to replicate, not knowing what you set.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Evaluation

\chapter{Evaluation}

\emph{Show graphs, run by run etc.}

\emph{Variance is interesting to discuss}

\emph{Could also be interesting to query what happens in an elbow in the graph - visual and a bit subjective. Three categories - those that don't take off, those that suddenly take off and those that gradually take off.}

\emph{Variance in each run, consequences of initialisation/activation, bigger questions about how reliable neural networks are, whether they're a good model for language evolution. }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Conclusion

\chapter{Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% the bibliography
\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{refs}
\bibliographystyle{apalike}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% the appendices
\appendix

\chapter{Project Proposal}

\input{proposal}

\end{document}
